---
title: "CAnD3 Replication Project"
author: "Jordan Katz"
date: "`r Sys.Date()`"
output: html_document
---


```{r setup, include=FALSE, warning=FALSE, echo=FALSE}

# install and load packages

# the packages for this code file can be installed using the code in lines 13-18

#install.packages("tidyverse")
#install.packages("modelsummary")
#install.packages("kableExtra")
#install.packages("gt")
#install.packages("broom")
#install.packages("MASS")

library(knitr)
library(tidyverse)
library(modelsummary)
library(kableExtra)
library(gt)
library(broom)
library(MASS)

```



```{r, include=FALSE, warning=FALSE, echo=FALSE}

# load data

# replace the file path in the setwd() function with your working directory. 

# set working directory
setwd("/Users/jkatz1/Documents/McGill/CAnD3/Replication Assignment")

# to import data, unzip the folder "GSS 2017 - Family, Cycle 31"
gss <- read.csv("./Replication Materials/GSS 2017 - Family, Cycle 31/gss-12M0025-E-2017-c-31_F1.csv")


```


### Data cleaning for analysis

- We recode the mental health variable (SRH_115) for easier interpretation. Lower values correspond to lower self-rated mental health. Higher values correspond to higher self-rated mental health.  

- We also recode each of the following independent variables to binary variables: SEX, MARSTAT, and BRTHCAN. These variables are now coded as flags for *female*, *married*, and *migrant*. In addition, we include a continuous variable for *age*. 

- All respondent skips and "don't know" responses are coded as NA and excluded from the analysis.  

```{r, include=FALSE, warning=FALSE, echo=FALSE}

gss_recode <- gss %>% 
  # select variables for analysis
  dplyr::select(SRH_115, BRTHCAN, MARSTAT, AGEC, SEX) %>%
  mutate(
    # recode SRH_115 variable for more intuitive interpretation 
    mental.health = case_when(
      SRH_115 == 1 ~ 5,
      SRH_115 == 2 ~ 4, 
      SRH_115 == 3 ~ 3,
      SRH_115 == 4 ~ 2, 
      SRH_115 == 5 ~ 1, 
      TRUE ~ NA),
    # create binary marital status variale
    married = case_when(
      MARSTAT == 1 ~ 1, 
      MARSTAT >= 2 & MARSTAT <= 6 ~ 0,
      TRUE ~ NA), 
    # recode SEX variable to make binary female variable
    female = case_when(
      SEX == 1 ~ 0,
      SEX == 2 ~ 1, 
      TRUE ~ NA),
    # drop people who did not respond to age question
    age = if_else(AGEC >= 96, NA, AGEC),
    # recode BRTHCAN variable to make binary migrant flag
    migrant = case_when(
      BRTHCAN == 1 ~ 0, 
      BRTHCAN == 2 ~ 1, 
      TRUE ~ NA)
    ) 

```

### Descriptive statistics

- Here, we include summary statistics (mean, standard deviation, and respondent count) of the variables used in the below models. 

```{r, warning=FALSE, echo=FALSE}

# calculate mean, standard deviation, and count
descriptives <- gss_recode %>%
  summarise(across(c(mental.health:migrant), list(
    mean = ~mean(.x, na.rm = TRUE),
    sd = ~sd(.x, na.rm = TRUE),
    n = ~sum(!is.na(.))
  ), .names = "{col}_{fn}"))

# reshape summary statistics
descriptives_long <- descriptives %>%
  pivot_longer(cols = everything(), names_to = c("variable", ".value"), names_sep = "_") 
 
# convert descriptives to table
descriptives_long %>%
  kable("html", digits = 2, caption = "Descriptive Statistics",
        col.names = c("Variable", "Mean", "Stan Dev", "N")) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    font_size = 12,
    position = "center",
    full_width = FALSE)

```


### Ordinal logistic regression models 

- We run two models---a basic model and a full model. The basic model (model 1) regresses self-rated mental health on the indicator variable for respondents not born in Canada (first generation migrants). The full model (model 2) introduces the other independent variables: *female*, *married*, *migrant*, and *age*. 

```{r, warning=FALSE, echo=FALSE, results='asis'}

# basic model that just looks at migrant variable 
model_basic <- polr(as.factor(mental.health) ~ migrant, data = gss_recode, Hess = TRUE)

# summary(model_basic)

# basic model that just looks at migrant variable 
model_full <- polr(as.factor(mental.health) ~ migrant + married + female + age,
                   data = gss_recode, Hess = TRUE)

# summary(model_full)

# combine the models into a list for results table 
models <- list(
  "Model 1" = model_basic,
  "Model 2" = model_full)

# create a summary table
model_summary <- modelsummary(models, stars = TRUE, output = "kableExtra",
                              coef_map = c("migrant", "married", "female", "age", 
                                           "1|2", "2|3", "3|4", "4|5"), 
                              coef_omit = "^$") # ensures order of coefficients

# Print the table with kableExtra styling
model_summary %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    font_size = 12,
    position = "center") 

```


### Results

- The analysis suggests that being married and having been born outside of Canada are correlated with higher levels of self-rated mental health. 

- In contrast, we see that females and aging are associated with lower levels of self-reported mental health. 


